---
title: "Plotly and `flexdashboard`"
output:
  html_document: 
    toc: true
    toc_float: true
---

In this page, we introduce `flexdashboards` as a way to collect plots (either static or interactive).


```{r, include = FALSE,message=FALSE,warning=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Slide Deck

<div class="vid_container">
  <iframe 
    src="https://speakerdeck.com/player/d593b18900444bf48c2516233f961139"
    allowfullscreen 
    frameborder="0"
    class="video">
  </iframe>
</div>

<div style="margin-bottom:5px"> <strong> <a href="https://speakerdeck.com/jeffgoldsmith/aphrea-workshop-dashboards" title="Dashboards" target="_blank">Plotly and Dashboards</a> </strong> from <strong><a href="https://speakerdeck.com/jeffgoldsmith" target="_blank">Jeff Goldsmith</a></strong>. </div><br>

## Example

### `flexdashboard`

Clearly you can embed interactive graphics in HTML files produced by R Markdown; this is a handy time to introduce dashboards. In short, dashboards are a collection of related graphics (or tables, or other outputs) that are displayed in a structured way that's easy to navigate. 

You can create dashboards using the `flexdashboard` package by specifying `flex_dashboard` as the output format in your R Markdown YAML. There are a variety of layout options, but we'll focus on a pretty simple structure produced by the template below. This is the default dashboard template in R Studio -- if you have `flexdashboard` installed, you can use `File > New File > R Markdown > From Template` to create a new .Rmd file with the structure below.

```{r, echo = FALSE, comment = ""}
cat(htmltools::includeText("resources/dashboard_template.Rmd"))
```

This dashboard has space for three plots, which we'll populate with plots of a few different types. Dashboard layouts are controlled by specifying columns and rows, and potentially subdiving these. We specified a two-column layout with set column widths, and then divided the second column into two panels. Using tabbed browsing and multiple pages can also be really useful -- check out the gallery linked below for examples!

### Load packages and data

I'll start by loading packages used in to create the dashboard. 

```{r}
library(tidyverse)
library(p8105.datasets)
```

We're going to focus on the [Airbnb](dataset_airbnb.html) data for this page. The code below extracts what we need right now; specifically, we select only a few of the variables and filter to include a subset of the data. In part, this makes sure that the resulting dataset and plots are computationally feasible -- for large datasets, you may need to downsample.

```{r import_data}
data(nyc_airbnb)

nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(rating = review_scores_location / 2) %>%
  rename(latitude = lat, longitude = long) %>%
  select(
    borough = neighbourhood_group, neighbourhood, rating, price, room_type,
    latitude, longitude) %>%
  filter(
    !is.na(rating), 
    borough == "Manhattan",
    room_type == "Entire home/apt",
    price %in% 100:500)
```

We'll use this dataset as the basis for our plots. 

### Plots

The three code chunks below make plots exploring this dataset. Right now, we're primarily interested in being able to create plots and embed them in a dashboard, so the plots themselves are relatively simple. 

Below we're plotting the location (latitude and longitude) of the rentals in our dataset, and mapping `price` to color. 

```{r}
nyc_airbnb %>%
  ggplot(aes(x = latitude, y = longitude, color = price)) + 
  geom_point(alpha = .5)
```

Next we show a boxplot for the distribution of price in neighbourhoods.

```{r}
nyc_airbnb %>% 
  mutate(neighbourhood = fct_reorder(neighbourhood, price)) %>% 
  ggplot(aes(x = neighbourhood, y = price)) + 
  geom_boxplot()
```

Lastly, we make a bar chart showing the number of rentals in each neighbourhood. 

```{r}
nyc_airbnb %>% 
  count(neighbourhood) %>% 
  mutate(neighbourhood = fct_reorder(neighbourhood, n)) %>% 
  ggplot(aes(x = neighbourhood, y = n)) + 
  geom_bar(stat = "identity")
```

Embedding these in the "open" spots in the flexdashboard will create a (static) way of visualizing these data. 

### `flexdashboard`s on websites

You can share the HTML files for dashboards directly (e.g. by email); you can also host these online to make the dashboard visible to others. That process is essentially the same as for any other website you'd make. 

However, the website's `_site.yml` file conflicts with the dashboard's `YAML` header regarding the output format -- and the website's `_site.yml` "wins". To address this issue, instead of knitting you can use this command to knit the dashboard.

```{r, eval = FALSE}
rmarkdown::render("dashboard_template.Rmd", output_format = "flexdashboard::flex_dashboard")
```

This will create `dashboard_template.html` but not open it in RStudio's Viewer pane; you can open the file in a browser instead. Alternatively, using RStudio's Build pane to Build Website will produce the same results. To illustrate, we'll put the dashboard we just created on a website for this topic. 

All of this `YAML` business is only an issue for dashboards embedded in websites; a standalone dashboard (in a non-website GH repo or R Project) can be knit using the same process as other .Rmd files. 


## Other materials

* Dashboards are pretty well-supported. Check out the [overview](https://pkgs.rstudio.com/flexdashboard/articles/flexdashboard.html), [layout](https://pkgs.rstudio.com/flexdashboard/articles/layouts.html) discussion, and [examples](https://pkgs.rstudio.com/flexdashboard/articles/examples.html)
* There are cool dashboards all over. To get a sense of how these work in the real world, check out: 
    * The NYC Restaurant Inspection [dashboard](https://data.cityofnewyork.us/Health/DOHMH-New-York-City-Restaurant-Inspection-Results/rs6k-p7g6);
    * This [app](https://juliasilge.shinyapps.io/power-app/) for understanding power (blog post [here](https://juliasilge.com/blog/ab-testing/));
    * 538's [p-hacking tool](https://projects.fivethirtyeight.com/p-hacking/index.html?initialWidth=1024&childId=phacking&parentTitle=Science%20Isn’t%20Broken%20%7C%20FiveThirtyEight&parentUrl=https%3A%2F%2Ffivethirtyeight.com%2Ffeatures%2Fscience-isnt-broken%2F) (full article [here](https://fivethirtyeight.com/features/science-isnt-broken/#part3)).

